$K$-corrections and filter transformations \\ in the ultraviolet, optical, and near infrared

### 2006-08-21 docs/paper/kcorrect.tex
add
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@565 75dff457-8751-0410-ab1a-cd701654ce6b

\documentclass[10pt,preprint]{aastex}
\usepackage{amsmath}
\usepackage{epsfig}

@@

\def\beq#1\eeq{\begin{equation}#1\end{equation}}
\def\beqa#1\eeqa{\begin{align}#1\end{align}}
\def\beqq#1\eeqq{\begin{equation*}#1\end{equation*}}
\def\beqqa#1\eeqqa{\begin{align*}#1\end{align*}}
\def\bseq#1\eseq{\begin{subequations}#1\end{subequations}}

@@

\subsection{Standard NMF}
The standard non-negative matrix factorization (NMF) problem, as
originally posed by \cite{lee00a}, is to approximate a data matrix
$\XX$ (of size $N_k\times N_n$) as the outer product of two rank-$N_i$
non-negative matrices $\WW$ and $\HH$. For example, in this paper,
$\XX$ represents the fluxes at $N_n$ different wavelengths for $N_k$
different galaxies, $\HH$ represents a set of $N_i$ templates to build
each galaxy from, and $\WW$ represents, for each galaxy, the weights
to give each template. This factorization is similar in spirit to
singular value decomposition (SVD) and PCA, but crucially, in NMF,
$\XX$, $\WW$ and $\HH$ are all constrained to have non-negative
entries. This changes the problem fundamentally, since no
``cancellation'' of positive and negative basis functions or
coefficients is possible -- all approximation interactions are
strictly additive. The goal, as with SVD or PCA, is to minmize the
squared approximation error (Frobenius norm): \beqa \label{eq:nmfcost}
\chi^2 &= \|\XX-\WW\HH\|^2\\
%&= \|\XX-\XXh\|^2\\
&= \sum_{kn} \left(\Xkn - \sum_i \Wki\Hin \right)^2
\eeqa
In their seminal papers, \cite{lee00a} show that this approximation
error in non-increasing under the following very simple multiplicative
update rules: 
\beqa \label{eq:wupdate}
\Wki &\leftarrow \Wki \frac{[\XX\HH\T]_{ki}}{[\WW\HH\HH\T]_{ki}}\\
\label{eq:hupdate}
\Hin &\leftarrow \Hin \frac{[\WW\T\XX]_{in}}{[\WW\T\WW\HH]_{in}}
\eeqa
$[M]_{ab}$ denotes the $(ab)$ element of the matrix valued expression
$M$. $\HH\T$ is the transpose of matrix $\HH$.

@@

\subsection{NMF in the space of coefficients for a known basis}
The factorization problem we face in this paper is a slight variation on the
basic NMF setup described above. We wish to approximate $\XX$ by the
product $\AA\BB\MM$, 
where $\MM$ is a given (fixed) basis matrix of size $N_j \times N_n$
and the optimization is over the matrices $\AA$ (of size
$N_k\times N_i$) and $\BB$ (of size $N_i\times N_j$). All matrices
$\XX,\AA,\BB\,\MM$ have non-negative entries. This can be thought of
as performing non-negative matrix factorization on the coefficients of
an approximation, given a fixed (non-negative) basis, described by the
columns of $\MM$. Once again, the objective we wish to minimize is the
squared approximation error: 
\beqa 
\chi^2 &= \|\XX-\AA\BB\MM\|^2\\
&= \sum_{kn} \left(\Xkn - \sum_{ij} \Aki\Bij\Mjn \right)^2
\eeqa
This equation is almost that posed in Equation \ref{chi2}.
In the case of this paper, this generalization allows us to express
the templates in terms of star-formation histories, but to compare the
predicted fluxes for the star-formation histories through the $\MM$
matrix to the observed fluxes $\XX$.

@@

First, by thinking of the tensor product $\BB\MM$ as a single
non-negative matrix $\HH$, we can trivially derive a multiplicative
update equation for the elements of $\AA$ by using the $\WW$ update
provided above: 
\beq
\Aki \leftarrow \Aki 
\frac{[\XX\MM\T\BB\T]_{ki}}{[\AA\BB\MM\MM\T\BB\T]_{ki}}
\eeq
By Lee and Seung's original proof, this update is guaranteed (for any
non-negative matrices $\XX$,$\MM$ and $\BB$) not to increase the
approximation error.

Our main algorithmic contribution is to derive a similar update equation
for the elements of $\BB$:
\beq \label{eq:bupdate}
\Bij \leftarrow \Bij
\frac{[\AA\T\XX\MM\T]_{ij}}{[\AA\T\AA\BB\MM\MM\T]_{ij}}
\eeq
Following the technique outlined in \cite{lee00a}, it can still be shown
that the error is nonincreasing under the application of this update.

@@

{\bf Lemma:} 
{\it For any symmetric matrix $\PP$ having non-negative entries $\Pni
\geq 0$, any vector $\zz$ having non-negative entries $\zn \geq 0$,
and any vector $\yy$,}
\beq
\sum_{ni} \yn \yi \Pni \leq \sum_{ni} \yn^2 \frac{\zi}{\zn} \Pni 
\eeq
%{\it If the vector $\yy$ has non-negative entries, equality can be
%achieved by choosing $\zz=\yy$.}

For the standard NMF problem, we can prove that the approximation
error \eqref{eq:nmfcost} is non-increasing under \eqref{eq:wupdate} by
using the above lemma to construct a function $\phi(\WW,\ZZ)$ which is
an upper bound on the cost $\chi^2(\WW)$ for any non-negative matrix $\ZZ$:
{\small 
\beqa 
\chi^2(\WW) &= \sum_{kni\ell} \Wki\Wkl \Hin\Hln
- 2 \sum_{kni} \Wki \Xkn \Hin + \sum_{kn} \Xkn^2 \\
\phi(\WW,\ZZ) &= \sum_{kni\ell} \Wki^2\frac{\Zkl}{\Zki} \Hin\Hln 
- 2 \sum_{kni} \Wki \Xkn \Hin + \sum_{kn} \Xkn^2 \\
\phi(\WW,\ZZ) &\geq \chi^2(\WW) \qquad \forall \: \Wki\geq 0,\: \Zki\geq 0
\eeqa
} with equality being achieved when $\ZZ=\WW$. Figure \ref{lemma}
represents this definition schematically, showing the true $\chi^2$
function and $\phi$. The trick of the method is to define $\phi$ such
that it is equal to $\chi^2$ at $W$, greater than $\chi^2$ everywhere
else, and is easily minimizable. Then one can use the minimum of
$\phi$ as the update, and be guaranteed that one's updates do not
increase $\chi^2$.

The function $\phi$ can be analytically minimized with respect to
its first argument:
\beqa
\phi(\WW^*,\ZZ) &\leq \phi(\WW,\ZZ) \qquad \forall \: \WW,\ZZ \\
\Wki^* &= \Zki \frac{\sum_n \Xkn \Hin}{\sum_{n\ell} \Hin\Hln \Zkl}
\eeqa
If we set $\ZZ=\WW$, this becomes exactly \eqref{eq:wupdate} and
now we can easily prove the validity of this update rule: 
\beq
\chi^2(\WW) = \phi(\WW,\WW) \geq \phi(\WW^*,\WW) \geq \chi^2(\WW^*)
\eeq
where the first inequality comes from the fact that $\WW^*$ minimizes
$\phi(\cdot,\WW)$ with respect to its first argument
and the second comes from the fact that $\phi(\WW^*,\cdot)$ is a bound
on $\chi^2(\WW^*)$. The proof of validity for the update
\eqref{eq:hupdate} is analogous by symmetry.

To prove the validity of our update \eqref{eq:bupdate} for our new
problem, we proceed in a similar fashion, using the lemma twice to
construct consecutive upper bounds on the cost:
{\small 
\beqa 
\chi^2(\BB) &= \sum_{kni\ell{}js} \Aki\Akl\Bij\Bls\Mjn\Msn\\ \nonumber
&- 2 \sum_{knij} \Xkn\Aki\Bij\Mjn + \sum_{kn} \Xkn^2 \\
\phi(\BB,\ZZ,\beta) &= \sum_{kni\ell{}js} \Aki\Akl\Bij^2
\frac{\Zks\betalj}{\Zkj\betaij}\Mjn\Msn\\ \nonumber
&- 2 \sum_{knij} \Xkn\Aki\Bij\Mjn + \sum_{kn} \Xkn^2 \\
\phi(\BB,\ZZ,\beta) &\geq \chi^2(\BB) \qquad \forall 
\: \Bij\geq 0,\: \Zkj\geq 0,\: \betaij\geq 0
\eeqa
} with equality being achieved when $\ZZ=\AA\BB$ and 
$\beta=\BB$.

Once again, the bound $\phi$ can be analytically minimized with respect to
its first argument:
\beqa
\phi(\BB^*,\ZZ,\beta) &\leq \phi(\BB,\ZZ,\beta) \qquad 
\forall \: \BB,\ZZ,\beta \\
\Bij^* &= \betaij \frac{\sum_{kn} \Xkn\Aki\Mjn}
{\sum_{kn\ell{}s} \Aki\Akl\frac{\Zks\betalj}{\Zkj}\Mjn\Msn}
\eeqa
If we set $\ZZ=\AA\BB$ and $\beta=\BB$, this becomes
exactly \eqref{eq:bupdate} and 
we can now prove the validity of this update rule: 
\beq
\chi^2(\BB) = \phi(\BB,\AA\BB,\BB) \geq \phi(\BB^*,\AA\BB,\BB) 
\geq \chi^2(\BB^*) 
\eeq
where the first inequality comes from the fact that $\BB^*$ minimizes
$\phi(\cdot,\AA\BB,\BB)$ with respect to its first argument
and the second comes from the fact that $\phi(\BB^*,\cdot,\cdot)$ is a bound
on $\chi^2(\BB^*)$. The only condition we require is that the elements
of the matrix $\MM\MM\T$ are all non-negative.

\subsection{Nonuniform uncertainties}
When different entries $\Xkn$ of the data matrix have different
uncertainties, the natural objective function is the weighted approximation
error (which also corresponds to the negative log likelihood under a
Gaussian noise assumption):
\beq
\chi^2 = \sum_{kn} \left( \frac{\Xkn - 
\sum_i \Wki\Hin}{\skn}\right)^2
\eeq

This case can easily be handled since during the update for $\WW$ the
elements of $\HH$ are fixed and vice versa and the updates are
guaranteed not to increase the cost for any nonnegative matrices.
%In particular, define $\XXt_{ij}=\Xkn/\skn$,
%$\WWt(j)_{ik}=\Wki/\skn$ and 
%$\HHt(i)_{kj}=\Hin/\skn$. 
In particular, when updating $\WW$, we can rewrite the cost function as
\beq
\chi^2 = \sum_{kn} \left( \frac{\Xkn}{\skn} - 
\sum_i \Wki\frac{\Hin}{\skn}\right)^2
%= \sum_{ij} \left(\XXt_{ij} - 
%\sum_k \Wki\HHt(i)_{kj}\right)^2
\eeq
yielding updates of the form:
\beq
\Wki \leftarrow \Wki 
\left( \sum_n \frac{\Xkn\Hin}{\skn^2} \right) / 
\left( \sum_{mn} \frac{W_{km}H_{mn}\Hin}{\skn^2}\right)
%\Wki \leftarrow \Wki \frac{[\XXt\HHt(i)\T]_{ik}}{[\WW\HHt(i)\HHt(i)\T]_{ik}}
\eeq

and similarly, when updating $\HH$, we can rewrite the cost function as
\beq
\chi^2 = \sum_{kn} \left( \frac{\Xkn}{\skn} - 
\sum_i \frac{\Wki}{\skn}\Hin\right)^2
%= \sum_{ij} \left(\XXt_{ij} - 
%\sum_k \WWt(j)_{ik}\Hin\right)^2
\eeq
yielding updates of the form:
\beq
\Hin \leftarrow \Hin
\left( \sum_k \frac{\Wki\Xkn}{\skn^2}\right) /
\left( \sum_{mk} \frac{\Wki{}W_{km}H_{mn}}{\skn^2}\right)
%\Hin \leftarrow \Hin \frac{[\WWt(j)\T\XX]_{kj}}{[\WWt(j)\T\WWt(j)\HH]_{kj}}
\eeq

This argument can be equally applied to our extended model, yielding
the final update equations which are actually implemented in the
kcorrect code:
%%\centerline{\fbox{\parbox{4.3in}{
\beqa 
\Aki &\leftarrow \Aki 
\left( \sum_{jn} \frac{\Xkn\Mjn\Bij}{\skn^2}\right) / 
\left( \sum_{mljn} \frac{A_{km}B_{mj}M_{jn}M_{nl}B_{il}}{\skn^2}\right) \\
\Bij &\leftarrow \Bij
\left( \sum_{kn} \frac{\Aki\Xkn\Mjn}{\skn^2}\right) / 
\left( \sum_{mlkn} \frac{A_{ki}A_{km}B_{ml}M_{ln}M_{jn}}{\skn^2}\right)
\eeqa
%%}}}

>>>>

\documentclass[10pt,preprint]{aastex}

@@

@@

\subsection{Standard NMF}
The standard non-negative matrix factorization (NMF) problem, as
originally posed by \cite{lee00a}, is to approximate a data matrix
$\XX$ (of size $N_k\times N_n$) as the outer product of two rank-$N_i$
non-negative matrices $\WW$ and $\HH$. For example, in this paper,
$\XX$ represents the fluxes at $N_n$ different wavelengths for $N_k$
different galaxies, $\HH$ represents a set of $N_i$ templates to build
each galaxy from, and $\WW$ represents, for each galaxy, the weights
to give each template. This factorization is similar in spirit to
singular value decomposition (SVD) and PCA, but crucially, in NMF,
$\XX$, $\WW$ and $\HH$ are all constrained to have non-negative
entries. This changes the problem fundamentally, since no
``cancellation'' of positive and negative basis functions or
coefficients is possible -- all approximation interactions are
strictly additive. The goal, as with SVD or PCA, is to minmize the
squared approximation error (Frobenius norm): 
\begin{eqnarray}
\label{eq:nmfcost}
\chi^2 &=& \|\XX-\WW\HH\|^2\\
%&= \|\XX-\XXh\|^2\\
&=& \sum_{kn} \left(\Xkn - \sum_i \Wki\Hin \right)^2
\end{eqnarray}
In their seminal papers, \cite{lee00a} show that this approximation
error in non-increasing under the following very simple multiplicative
update rules: 
\begin{eqnarray}
 \label{eq:wupdate}
\Wki &\leftarrow& \Wki \frac{[\XX\HH\T]_{ki}}{[\WW\HH\HH\T]_{ki}}\\
\label{eq:hupdate}
\Hin &\leftarrow& \Hin \frac{[\WW\T\XX]_{in}}{[\WW\T\WW\HH]_{in}}
\end{eqnarray}
$[M]_{ab}$ denotes the $(ab)$ element of the matrix valued expression
$M$. $\HH\T$ is the transpose of matrix $\HH$.

@@

\subsection{NMF in the space of coefficients for a known basis}
The factorization problem we face in this paper is a slight variation on the
basic NMF setup described above. We wish to approximate $\XX$ by the
product $\AA\BB\MM$, 
where $\MM$ is a given (fixed) basis matrix of size $N_j \times N_n$
and the optimization is over the matrices $\AA$ (of size
$N_k\times N_i$) and $\BB$ (of size $N_i\times N_j$). All matrices
$\XX,\AA,\BB\,\MM$ have non-negative entries. This can be thought of
as performing non-negative matrix factorization on the coefficients of
an approximation, given a fixed (non-negative) basis, described by the
columns of $\MM$. Once again, the objective we wish to minimize is the
squared approximation error: 
\begin{eqnarray} 
\chi^2 &=& \|\XX-\AA\BB\MM\|^2\\
&=& \sum_{kn} \left(\Xkn - \sum_{ij} \Aki\Bij\Mjn \right)^2
\end{eqnarray}
This equation is almost that posed in Equation \ref{chi2}.
In the case of this paper, this generalization allows us to express
the templates in terms of star-formation histories, but to compare the
predicted fluxes for the star-formation histories through the $\MM$
matrix to the observed fluxes $\XX$.

@@

First, by thinking of the tensor product $\BB\MM$ as a single
non-negative matrix $\HH$, we can trivially derive a multiplicative
update equation for the elements of $\AA$ by using the $\WW$ update
provided above: 
\begin{equation}
\Aki \leftarrow \Aki 
\frac{[\XX\MM\T\BB\T]_{ki}}{[\AA\BB\MM\MM\T\BB\T]_{ki}}
\end{equation}
By Lee and Seung's original proof, this update is guaranteed (for any
non-negative matrices $\XX$,$\MM$ and $\BB$) not to increase the
approximation error.

Our main algorithmic contribution is to derive a similar update equation
for the elements of $\BB$:
\begin{equation} \label{eq:bupdate}
\Bij \leftarrow \Bij
\frac{[\AA\T\XX\MM\T]_{ij}}{[\AA\T\AA\BB\MM\MM\T]_{ij}}
\end{equation}
Following the technique outlined in \cite{lee00a}, it can still be shown
that the error is nonincreasing under the application of this update.

@@

{\bf Lemma:} 
{\it For any symmetric matrix $\PP$ having non-negative entries $\Pni
\geq 0$, any vector $\zz$ having non-negative entries $\zn \geq 0$,
and any vector $\yy$,}
\begin{equation}
\sum_{ni} \yn \yi \Pni \leq \sum_{ni} \yn^2 \frac{\zi}{\zn} \Pni 
\end{equation}
%{\it If the vector $\yy$ has non-negative entries, equality can be
%achieved by choosing $\zz=\yy$.}

For the standard NMF problem, we can prove that the approximation
error (Equation \ref{eq:nmfcost}) is non-increasing under Equation
\ref{eq:wupdate} by using the above lemma to construct a function
$\phi(\WW,\ZZ)$ which is an upper bound on the cost $\chi^2(\WW)$ for
any non-negative matrix $\ZZ$: {\small
\begin{eqnarray} 
\chi^2(\WW) &=& \sum_{kni\ell} \Wki\Wkl \Hin\Hln
- 2 \sum_{kni} \Wki \Xkn \Hin + \sum_{kn} \Xkn^2 \\
\phi(\WW,\ZZ) &=& \sum_{kni\ell} \Wki^2\frac{\Zkl}{\Zki} \Hin\Hln 
- 2 \sum_{kni} \Wki \Xkn \Hin + \sum_{kn} \Xkn^2 \\
\phi(\WW,\ZZ) 
&&\geq \chi^2(\WW) \qquad \forall \: \Wki\geq 0,\: \Zki\geq 0
\end{eqnarray}
} with equality being achieved when $\ZZ=\WW$. Figure \ref{lemma}
represents this definition schematically, showing the true $\chi^2$
function and $\phi$. The trick of the method is to define $\phi$ such
that it is equal to $\chi^2$ at $W$, greater than $\chi^2$ everywhere
else, and is easily minimizable. Then one can use the minimum of
$\phi$ as the update, and be guaranteed that one's updates do not
increase $\chi^2$.

The function $\phi$ can be analytically minimized with respect to
its first argument:
\begin{eqnarray}
\phi(\WW^*,\ZZ) &\leq& \phi(\WW,\ZZ) \qquad \forall \: \WW,\ZZ \\
\Wki^* &=& \Zki \frac{\sum_n \Xkn \Hin}{\sum_{n\ell} \Hin\Hln \Zkl}
\end{eqnarray}
If we set $\ZZ=\WW$, this becomes exactly Equation \ref{eq:wupdate} and
now we can easily prove the validity of this update rule: 
\begin{equation}
\chi^2(\WW) = \phi(\WW,\WW) \geq \phi(\WW^*,\WW) \geq \chi^2(\WW^*)
\end{equation}
where the first inequality comes from the fact that $\WW^*$ minimizes
$\phi(\cdot,\WW)$ with respect to its first argument
and the second comes from the fact that $\phi(\WW^*,\cdot)$ is a bound
on $\chi^2(\WW^*)$. The proof of validity for the update
(Equation \ref{eq:hupdate}) is analogous by symmetry.

To prove the validity of our update (Equation \ref{eq:bupdate}) for
our new problem, we proceed in a similar fashion, using the lemma
twice to construct consecutive upper bounds on the cost: {\small
\begin{eqnarray} 
\chi^2(\BB) &=& \sum_{kni\ell{}js} \Aki\Akl\Bij\Bls\Mjn\Msn\\ \nonumber
&&- 2 \sum_{knij} \Xkn\Aki\Bij\Mjn + \sum_{kn} \Xkn^2 \\
\phi(\BB,\ZZ,\beta) &=& \sum_{kni\ell{}js} \Aki\Akl\Bij^2
\frac{\Zks\betalj}{\Zkj\betaij}\Mjn\Msn\\ \nonumber
&&- 2 \sum_{knij} \Xkn\Aki\Bij\Mjn + \sum_{kn} \Xkn^2 \\
\phi(\BB,\ZZ,\beta) &&\geq \chi^2(\BB) \qquad \forall 
\: \Bij\geq 0,\: \Zkj\geq 0,\: \betaij\geq 0
\end{eqnarray}
} with equality being achieved when $\ZZ=\AA\BB$ and 
$\beta=\BB$.

Once again, the bound $\phi$ can be analytically minimized with respect to
its first argument:
\begin{eqnarray}
\phi(\BB^*,\ZZ,\beta) &\leq& \phi(\BB,\ZZ,\beta) \qquad 
\forall \: \BB,\ZZ,\beta \\
\Bij^* &=& \betaij \frac{\sum_{kn} \Xkn\Aki\Mjn}
{\sum_{kn\ell{}s} \Aki\Akl\frac{\Zks\betalj}{\Zkj}\Mjn\Msn}
\end{eqnarray}
If we set $\ZZ=\AA\BB$ and $\beta=\BB$, this becomes
exactly Equation \ref{eq:bupdate} and 
we can now prove the validity of this update rule: 
\begin{equation}
\chi^2(\BB) = \phi(\BB,\AA\BB,\BB) \geq \phi(\BB^*,\AA\BB,\BB) 
\geq \chi^2(\BB^*) 
\end{equation}
where the first inequality comes from the fact that $\BB^*$ minimizes
$\phi(\cdot,\AA\BB,\BB)$ with respect to its first argument
and the second comes from the fact that $\phi(\BB^*,\cdot,\cdot)$ is a bound
on $\chi^2(\BB^*)$. The only condition we require is that the elements
of the matrix $\MM\MM\T$ are all non-negative.

\subsection{Nonuniform uncertainties}
When different entries $\Xkn$ of the data matrix have different
uncertainties, the natural objective function is the weighted approximation
error (which also corresponds to the negative log likelihood under a
Gaussian noise assumption):
\begin{equation}
\chi^2 = \sum_{kn} \left( \frac{\Xkn - 
\sum_i \Wki\Hin}{\skn}\right)^2
\end{equation}

This case can easily be handled since during the update for $\WW$ the
elements of $\HH$ are fixed and vice versa and the updates are
guaranteed not to increase the cost for any nonnegative matrices.
%In particular, define $\XXt_{ij}=\Xkn/\skn$,
%$\WWt(j)_{ik}=\Wki/\skn$ and 
%$\HHt(i)_{kj}=\Hin/\skn$. 
In particular, when updating $\WW$, we can rewrite the cost function as
\begin{equation}
\chi^2 = \sum_{kn} \left( \frac{\Xkn}{\skn} - 
\sum_i \Wki\frac{\Hin}{\skn}\right)^2
%= \sum_{ij} \left(\XXt_{ij} - 
%\sum_k \Wki\HHt(i)_{kj}\right)^2
\end{equation}
yielding updates of the form:
\begin{equation}
\Wki \leftarrow \Wki 
\left( \sum_n \frac{\Xkn\Hin}{\skn^2} \right) / 
\left( \sum_{mn} \frac{W_{km}H_{mn}\Hin}{\skn^2}\right)
%\Wki \leftarrow \Wki \frac{[\XXt\HHt(i)\T]_{ik}}{[\WW\HHt(i)\HHt(i)\T]_{ik}}
\end{equation}

and similarly, when updating $\HH$, we can rewrite the cost function as
\begin{equation}
\chi^2 = \sum_{kn} \left( \frac{\Xkn}{\skn} - 
\sum_i \frac{\Wki}{\skn}\Hin\right)^2
%= \sum_{ij} \left(\XXt_{ij} - 
%\sum_k \WWt(j)_{ik}\Hin\right)^2
\end{equation}
yielding updates of the form:
\begin{equation}
\Hin \leftarrow \Hin
\left( \sum_k \frac{\Wki\Xkn}{\skn^2}\right) /
\left( \sum_{mk} \frac{\Wki{}W_{km}H_{mn}}{\skn^2}\right)
%\Hin \leftarrow \Hin \frac{[\WWt(j)\T\XX]_{kj}}{[\WWt(j)\T\WWt(j)\HH]_{kj}}
\end{equation}

This argument can be equally applied to our extended model, yielding
the final update equations which are actually implemented in the
kcorrect code:
%%\centerline{\fbox{\parbox{4.3in}{
\begin{eqnarray} 
\Aki &\leftarrow& \Aki 
\left( \sum_{jn} \frac{\Xkn\Mjn\Bij}{\skn^2}\right) / 
\left( \sum_{mljn} \frac{A_{km}B_{mj}M_{jn}M_{nl}B_{il}}{\skn^2}\right) \\
\Bij &\leftarrow& \Bij
\left( \sum_{kn} \frac{\Aki\Xkn\Mjn}{\skn^2}\right) / 
\left( \sum_{mlkn} \frac{A_{ki}A_{km}B_{ml}M_{ln}M_{jn}}{\skn^2}\right)
\end{eqnarray}
%%}}}
